---
title: "Leslie Matrix Example"
author: "Erica bishop"
format: html
editor: visual
---

##Leslie matrix population model example

Modeling population changes with fly example from lecture

*note: quarto very similar to R markdown, render with knitR or Jupyter notebooks, better at multiple language interoperability*

```{r}
##practice making a matrix, for 3x3 matrix need to create all nine vectors. can be made all in one step or in two separate steps. One step shown below in nested function

insect_leslie <- matrix(c(0, 0, 600, 0.2, 0, 0, 0, 0.08, 0), 
                        nrow = 3, 
                        ncol = 3, 
                        byrow = TRUE)

insect_leslie

```

```{r}
##now create initial condition, create vector for values, then can multiply matrix by vectors to solve system

insect_y0 <- c(12000, 700, 500)

###insect_y0 is single column matrix, now can multiply by insect_leslie, can store outputs

insect_y1 <- insect_leslie %*% insect_y0

insect_y1
```

Now estimate populations at each life stage at y=1, y=2, and y=3. Doing individually but a for loop would do together
%*% is dot product operator

We now have y1 so use y1 as new initial conidtion for y2

```{r}

insect_y2 <- insect_leslie %*% insect_y1

insect_y2
```

```{r}
insect_y3 <- insect_leslie %*% insect_y2

insect_y3
```


